---
title: "Task on Modeling"
author: "Ignacio Almodóvar & Alejandra Estrada"
date: "3/22/2022"
output: pdf_document
---

```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
library(ggplot2)
library(h2o)
library(tidymodels)
```

```{r}
datas=read.csv2("drug200.csv",sep = ",",header = TRUE)
summary(datas)
datas %>% head(5)
```
As we can see, the summary says that all variables except Age are factors. However, analying the data the variable "Na_to_k" looks like a numeric variable so we must change it.

```{r}
datas[,5]  %<>% as.numeric() 
summary(datas)
```

We now plot the continuous variables to see if we can find any group evidences for the type of drug


```{r}
ggplot(data,aes(Age,Na_to_K,col=Drug)) + geom_point()
```

Beforehand there is not clear evidence for the differentiation in groups given the age and the Na_t_k. However, as can be seen, for the DrugY there is a clear bandwidth for Na_to_k being higher than 15.

We are now going to fitt a classification model using h2o package.

```{r,results=FALSE}
table(data$Drug)
h2o.init()
data_h2o=as.h2o(data)
resp_data="Drug"
pred_data=setdiff(names(data_h2o), resp_data)


setdiff(names(data_h2o), resp_data)
data_h2o[, resp_data] <- as.factor(data_h2o[, resp_data])

splits = h2o.splitFrame(data = data_h2o, ratios = 0.8, seed = 42)
train = splits[[1]]
test = splits[[2]]
# Run AutoML
aml_mul = h2o.automl(x = pred_data, y = resp_data, training_frame = train, leaderboard_frame = test,
                      include_algos = c("GLM", "XGBoost", "DeepLearning"),max_runtime_secs = 100,max_runtime_secs_per_model = 100, verbosity = NULL)


```


```{r}
lb_mul <- h2o.get_leaderboard(aml_mul)
print(lb_mul, n = nrow(lb_mul))
```


```{r}
pred_mul <- h2o.predict(object = aml_mul, newdata = test)
h2o.head(pred_mul)

labels_mul <- as.matrix(pred_mul$predict)
table(labels_mul, as.matrix(test$Drug))
```



## Tidymodels


```{r}
boot_data <- bootstraps(data, times = 10)
analysis(boot_data$splits[[1]] )%>% head()
```
### Parsnip

```{r}
data(datas,package="modeldata")
```

### Discrim

```{r}
library(discrim)
# Fit a Naive Bayes model (which is actually a kernel discriminant # analysis done by combining univariate kernel density estimators) 

summary(data)
data$Drug %<>% as.factor() 
nb_mod <- naive_Bayes() %>%
  set_engine("naivebayes") %>%
  fit(Drug ~ Na_to_K + Sex, data = data)
nb_mod
```

